# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
check_first_charc <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix #check if the first several characters of string is same as prefix.
}
drop_last <- function(x) {
  if (length(x) <= 1) return(NULL)#judge if the input has length greater than 1, if not return NULL.
  x[-length(x)]#if the input has length greater than 1, drop the last element in the input.
}
matchLength <- function(x, y) {
  rep(y, length.out = length(x))#replicate y according to the length of x
}


```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}

#rnorm samples from the univariate normal distribution and mvrnorm samples fromthe multivariate normal distribution. They could be made more consistent by having the same name.
```


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}
unlist(lapply(mtcars, function(x) sd(x) / mean(x)))

```


4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}
vapply(mtcars,sd,numeric(1))
```
    
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
vapply(mtcars[vapply(mtcars,is.numeric,logical(1))],sd,numeric(1))
```
    
